Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-05-13 text/back_end_KS.qmd
chicago revisions

\newpage
\pagenumbering{gobble}
\pagenumbering{arabic}

# Data and Methods

@@

In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.^[For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to. Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.

The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major). College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) [CITE]. From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S. Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) [CITE].[^5] Information about student socioeconomic status is limited. However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].

[^5]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select. Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”

@@

**Research design**. We utilize a multiple, quantitative case study design. Metropolitan areas are cases. Case studies deepen understanding of phenomena by comparing and contrasting across different settings \[CITE YIN 2001\]. Our research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects. Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics [MAKE THIS MORE ABOUT EXPLORING THE MODEL. THIS SENTENCE SOUNDS FANCY BUT DOESN'T REALLY SAY ANYTHING]. Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts. We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\]. Online Appendix XXX presents results for a larger number of metropolitan areas.

Case selection was informed by purposeful sampling to identify "information-rich" cases (Patton, 2002; Yin, 2009)[CITE] that can provide knowledge about variation across Geomarkets and over time. We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data. Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p. 11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia. To compare and contrast with classic urban model structures, we also identified alternative urban models such as bimodal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City). The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S. regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations. Related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2). We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.

@@

-->
```

# Findings

@@

@fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods. The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982[p. 12]. Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020. In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets. The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most densely populated Geomarket.

These modest population sgufts belie substantial change in the racial/ethnic composition within and across Geomarkets over time. @fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarket across the three census periods. In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore. In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population. By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander). For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs. Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie). @fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods. <!-- For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020). City of Chicago also experienced relatively modest declines in its Black population (37% in 1980 to 33% in 2020), alongside increases in its Hispanic (15% in 1980 to 26% in 2020) and Asian Pacific Islander populations (4% in 2000 to 6% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area. -->

@@

From 1980 to 2020, these demographic shifts have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by racial/ethnic and socioeconomic characteristics. The original Geomarket boundaries delineated significant racial and income segregation in the 1980s. This included Black, Hispanic, and Asian residents primarily in the City of Chicago Geomarket and White residents in Geomarkets forming the suburban ring of the metropolitan area. Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs. However, declining White populations, Hispanic and Asian population growth, and Black out-migration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining. The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents) while the the Evanston and North Shore Geomarkets remain predominantly White. While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains across Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs). [I THINK SOME DETAIL CAN BE REDUCED HERE; ALSO, I THINK YOU ARE BURYING THE LEDE THAT NEEDS TO BE EMPHASIZED: THERE IS SUBSTANTIAL RACIAL AND SES INEQUALITY BETWEEN GEOMARKETS AND CHICAGO ESPECIALLY IS POOREST AND MOST NON-WHITE]

@@

DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES!
@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area. At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets generally reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race). For example, more than 8,000 of the 16,485 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations. Similar to the larger metropolitan area, the North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students. Interestingly, while the City of Chicago is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-chicago-rq2-race includes nearly double the proportion of Hispanic (50%) prospects relative to the larger Geomarket resident population (26% Hispanic) in @fig-chicago-rq1-race with a lower proportion of White prospects (14%).

However, the racial/ethnic composition of prospect pools become less similar to the overall metropolitan area as PSAT scores increase. Notably, the shares of White and Asian prospects within Geomarkets increase, while Black and Hispanic prospects decrease. For instance, Asian prospects over representation in Geomarkets relative to the overall resident population at the 1310-1600 SAT score threshold ranges from a minimum of 4 percentage points in the North Shore Geomarket (17% of prospects versus 13% of residents) to a maximum of 21 percentage points in the Chain of Lakes (28% of prospects versus 7% residents) Geomarket. The City of Chicago Geomarket exhibits the largest underrepresentation of Black prospects (7% of prospects versus 33% of residents) while Chain of Lakes exhibits the largest underrepresentation of Hispanic prospects (7% of prospects versus 24% of residents). [LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN]

@@

@fig-chicago-rq2-highSAT-combo shows the magnitude of Geomarket contribution disparities increases at the highest SAT score range (1310-1600), partly driven by the relatively fewer number of first-generation college students included within these orders. Again, the City of Chicago Geomarket disproportionately contributes a disproportionate share of first-generation Asian college students (43% across both parental education categories) and first-generation Hispanic college students (52% across both parental education categories) in comparison to the overall distribution of first-generation college students within each racial/ethnic category (13% for Asian and 40% for Hispanic, respectively). Conversely, nearly all other Geomarkets contribute a disproportionately larger share of non-first generation college students across all racial/ethnic categories. -->
```

More than half (n=8,256) of all prospects whose contact information was purchased in these middle SAT score orders identified as White. @fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects. One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects. On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively. Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.

In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race. Among the nearly 1,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects. The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.

Appendix figures \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting lower-range (1020-1150) and high-range (1310-1600) SAT scores. Both figures demonstrate how representational Geomarket contributionsacross race/ethnicity, such as City of Chicago disproportionately contributing larger shares of Black and Hispanic prospects, remain consistent even at lower and higher SAT score thresholds. [REVISE THESE SENTENCES TO HAMMER THE LEDE -- IF WE EXCLUDE CITY OF CHICAGO GEOMAREKT, THE WAY THE ASU ORDER DOES, WE LOSE A DISPROPORTIONATE SHARE OF BLACK AND HISPANIC STUDENTS]

@@

Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area. @fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 12,981 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders. For instance, among all White prospects (n=8,155) whose contact information was purchased in this SAT score range, approximately 6% were first-generation college students whose parents did not attend college, 15% were first-generation college students whose parents attended some college but did not complete their degree, and 79% were not first-generation college students. @fig-chicago-rq2-midSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (10% and 9%, respectively). For first-generation college students whose parents attended some college but never completed degrees, only the South and Southwest Suburbs (25%) Geomarket contributed a substantially larger share of White prospects. In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students. [MY READ IS THAT THE PATTERNS FOR WHITE ARE NOT THE SAME AS THE PATTERNS FOR ASIAN AND LATINX -- SO I WOULDN'T SAY THESE PATTERNS PERSIST ACROSS OTHER RACIAL/ETHNIC CATEGORIES. I WOULD RECOMMEND STARTING WITH ASIAN AND WITH HISPANIC. THAT IS THE LEDE; ALSO, THIS IS WHERE YOU COULD REFER TO THE SIMILAR GRAPH FOR PHILADELPHIA AND SAY IT FOLLOWS THE SAME PATTERN; I THINK THIS IS THE CLIMAX AND COOLEST INSIGHT OF THE RESULTS]

>>>>

```{=tex}
\newpage
\pagenumbering{gobble}
\pagenumbering{arabic}
```
# Data and Methods

@@

In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.[^5] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to. Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.

[^5]: For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.

The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major). College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) \[CITE\]. From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S. Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].[^6] Information about student socioeconomic status is limited. However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].

[^6]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select. Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”

@@

**Research design**. We utilize a multiple, quantitative case study design. Metropolitan areas are cases. Case studies deepen understanding of phenomena by comparing and contrasting across different settings \[CITE YIN 2001\]. Our research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects. Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics \[MAKE THIS MORE ABOUT EXPLORING THE MODEL. THIS SENTENCE SOUNDS FANCY BUT DOESN'T REALLY SAY ANYTHING\]. Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts. We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\]. Online Appendix XXX presents results for a larger number of metropolitan areas.

Case selection was informed by purposeful sampling to identify "information-rich" cases (Patton, 2002; Yin, 2009)[CITE] that can provide knowledge about variation across Geomarkets and over time. We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data. Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area[s] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p. 11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia. To compare and contrast with classic urban model structures, we also identified alternative urban models such as bimodal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City). The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S. regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations. Related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2). We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.

@@

-->
```
# Findings

@@

@fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods. The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 [p. 12]. Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020. In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets. The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most densely populated Geomarket.

These modest population shifts lead to substantial change in the racial/ethnic composition within and across Geomarkets over time. @fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarket across the three census periods. In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore. In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population. By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander). For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs. Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie). @fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods. <!-- For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020). City of Chicago also experienced relatively modest declines in its Black population (37% in 1980 to 33% in 2020), alongside increases in its Hispanic (15% in 1980 to 26% in 2020) and Asian Pacific Islander populations (4% in 2000 to 6% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area. -->

@@

Demographic shifts overtime have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by substantial racial/ethnic and socioeconomic inequality between Geomarkets. The original Geomarket boundaries delineated significant racial and income segregation in the 1980s. This included Black, Hispanic, and Asian residents primarily in the City of Chicago Geomarket and White residents in Geomarkets forming the suburban ring of the metropolitan area. Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs. However, declining White populations, Hispanic and Asian population growth, and Black out-migration, as well as narrowing of education and poverty gaps within some Geomarkets, have contributed to a more stratified Geomarket dynamic in the metropolitan area with significant disparities remaining between Geomarkets. The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets now also include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents). While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains between Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).

<!--# [I THINK SOME DETAIL CAN BE REDUCED HERE; ALSO, I THINK YOU ARE BURYING THE LEDE THAT NEEDS TO BE EMPHASIZED: THERE IS SUBSTANTIAL RACIAL AND SES INEQUALITY BETWEEN GEOMARKETS AND CHICAGO ESPECIALLY IS POOREST AND MOST NON-WHITE ————-  DALLAS IS POOREST AND MOST NON-WHITE; NOT CHI] -->

@@

<!--# DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES! --> @fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles. At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race). For example, more than 8,000 of the 15,981 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations. The North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects. Unlike any other Geomarket, nearly nine of every 10 prospects from the City of Chicago Geomarket were Hispanic (50%), Black (30%), Asian (5%), or Multiracial (2%).

The City of Chicago Geomarket increasingly becomes the only Geomarket with substantial shares of Black, and Hispanic prospects as PSAT scores increase. All Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline. For instance, the City of Chicago Geomarket pool is 18% Asian, 7% Black, 21% Hispanic, 5% Multiracial, making it the only Geomarket that is not predominantly White (49%). Prospect pools from all other Geomarkets in this high PSAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.4% (North Shore) to 5% (South and Southwest Suburbs) of prospects and 6% (Northwest Suburbs) to 10% (South and Southwest Suburbs) of prospects, respectively. <!--# [LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN] -->

@@

@fig-chicago-rq2-highSAT-combo shows the magnitude of Geomarket contribution disparities increases at the highest SAT score range (1310-1600), partly driven by the relatively fewer number of first-generation college students included within these orders. Again, the City of Chicago Geomarket disproportionately contributes a disproportionate share of first-generation Asian college students (43% across both parental education categories) and first-generation Hispanic college students (52% across both parental education categories) in comparison to the overall distribution of first-generation college students within each racial/ethnic category (13% for Asian and 40% for Hispanic, respectively). Conversely, nearly all other Geomarkets contribute a disproportionately larger share of non-first generation college students across all racial/ethnic categories. -->
```
For instance, @fig-chicago-rq2-midSAT-race shows the nearly 1,000 Black students whose contact information was purchased in these middle SAT score orders were largely contributed to the overall pool by the City of Chicago and South and Southwest Suburbs. These two Geomarkets contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects. On the other hand, all other Geomarkets contributed substantially smaller shares of Black prospects. The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.

In contrast, these representational patterns by Geomarket are reversed for White and Asian students. @fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates proportional to their overall contributions to all included prospects (except City of Chicago). Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects) and Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospectq8s) contributing smaller shares relative to their overall contribution to the larger prospect pool.

<!--# One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects. On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively.  -->

<!--# MOVED THIS UP In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race. Among the nearly 1,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects. The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.  -->

These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Chicago metropolitan area using Geomarket filters that exclude the City of Chicago– as observed in some orders collected via our public requests – result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools. Appendix \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race show that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters. In both cases, the City of Chicago contributes disproportionately larger shares of Black and Hispanic prospects compared to all other Geomarkets. <!--# [REVISE THESE SENTENCES TO HAMMER THE LEDE -- IF WE EXCLUDE CITY OF CHICAGO GEOMAREKT, THE WAY THE ASU ORDER DOES, WE LOSE A DISPROPORTIONATE SHARE OF BLACK AND HISPANIC STUDENTS] -->

@@

Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area. @fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 12,981 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders. @fig-chicago-rq2-midSAT-combo shows the City of Chicago Geomarket consistently contributes a disproportionately larger share of first-generation college students whose parents did not attend college across all racial/ethnic categories. For instance, among all Asian prospects (n=1,341) whose contact information was purchased in this SAT score range, approximately 12% were first-generation college students whose parents did not attend college, 16% were first-generation college students whose parents attended some college but did not complete their degree, and 72% were not first-generation college students. @fig-chicago-rq2-midSAT-combo shows the City of Chicago disproportionately contributes 35% of all Asian first-generation college students whose parents did not attend college. For first-generation college students whose parents attended some college but never completed degrees, City of Chicago (22%), South and Southwest Suburbs (20%), and Evanston and Skokie (20%) Geomarkets contributed larger shares of Asian prospects.

Similar Geomarket contribution patterns are evident among Hispanic prospects. The City of Chicago Geomarket stands out as the only Geomarket contributing substantially larger shares of Hispanic first-generation college students across those whose parents did not attend college (46% versus 38% overall) and those whose parents attended some college but did not complete their degrees (29% versus 26% overall). Although Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and Western Surburbs Geomarkets.

Lastly, although Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and Western Surburbs Geomarkets. Similarly, Geomarkets such as Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs contribute a disproportionately larger share of White prospects who are not first-generation college students, whereas City of Chicago and South and Southwest Suburbs contribute larger shares of White first-generation college students.

<!--# OLD WHITE PROSPECT TEXT among all White prospects (n=8,155) whose contact information was purchased in this SAT score range, approximately 6% were first-generation college students whose parents did not attend college, 15% were first-generation college students whose parents attended some college but did not complete their degree, and 79% were not first-generation college students. @fig-chicago-rq2-midSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (10% and 9%, respectively). For first-generation college students whose parents attended some college but never completed degrees, only the South and Southwest Suburbs (25%) Geomarket contributed a substantially larger share of White prospects. In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students. -->
